Finch: Sparse and Structured Array Programming with Control Flow

### 2024-03-27 sections/eval.tex

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. SpGEMM focuses on efficiently computing the product of two sparse matrices to produce an output sparse matrix.
The inner-product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. 
Inner-product involves multiplying two matrices by taking dot products of corresponding rows and columns.
Outer-product computes the product of all pairs of elements from the input matrices.
More detailed explanation is available in Section 2.2 of ~\cite{zhang2021gamma}.
Figures ~\ref{spgemm_small} and ~\ref{spgemm_large} show SpGEMM results with Finch and TACO.
Note that inner-product is very inefficient unless sparse matrices are nearly dense because the cost of intersections that incurs a lot of redundant computations. For the output-product, TACO only support the dense output: TACO does not support outer product due to its lack of support for scatter into a 2D workspace. So, due to memory constraints, we are unable to run TACO's outer-product for large sparse matrices. So, for large sparse matrices, we only compared Gustavson's algorithms. 
As shown in Figure~\ref{spgemm_large}, Finch outperforms TACO with an average speed

@@

\help{6. explain the features of "datastructure-driven programming" that are used here in this problem. i.e. explicit temporary, bytemap format, transpositions}.
\begin{figure}
    \label{spgemm_small}
	\includegraphics[width=\linewidth]{spgemm_small_speedup_log_scale.png}
    \caption{A comparison of several matrix multiplication algorithms between Finch and Taco smaller matrices, ordered from small to big dimension. Note that inner products necessarily requires $O(n^2)$ work and taco's outer products format is dense. Finch can use a sparse outer products format and thus has an asymptotic advantage that becomes evident as the output dimensions grow.}
\end{figure}

\begin{figure}
\label{spgemm_large}
	\includegraphics[width=\linewidth]{spgemm_joel_speedup.png}
    \caption{A comparison of gustavson's algorithm between Finch and Taco on some larger matrices}
\end{figure}

>>>>

Sparse Matrix-Matrix Multiplication (SpGEMM) is a fundamental operation in scientific computing and data analytics. SpGEMM focuses on efficiently computing the product of two sparse matrices to produce an output sparse matrix.
The inner-product, outer product, and Gustavson's algorithm are three different approaches to performing SpGEMM. 
Inner-product involves multiplying two matrices by taking dot products of corresponding rows and columns.
Outer-product computes the product of all pairs of elements from the input matrices.
More detailed explanation is available in Section 2.2 of ~\cite{zhang2021gamma}.
Figures ~\ref{fig:spgemm_small} and ~\ref{fig:spgemm_large} show SpGEMM results with Finch and TACO.
Note that inner-product is very inefficient unless sparse matrices are nearly dense because the cost of intersections that incurs a lot of redundant traversals. For the output-product, TACO only supports the dense output: TACO does not support outer product due to its lack of support for scatter into a 2D workspace. So, due to memory constraints, we are unable to run TACO's outer-product for large sparse matrices. On the other hand, 
So, for large sparse matrices, we only compared Gustavson's algorithms.
\changwan{mention TACO's gustavson's algorithm is hard-wired}
As shown in Figure~\ref{fig:spgemm_small}, for inner-product, Finch achieves comparable performance with TACO, and for outer-product, the 
As shown in Figure~\ref{fig:spgemm_large}, for large sparse matrices, Finch outperforms TACO with an average speedup of 1.05.

@@

\help{6. explain the features of "datastructure-driven programming" that are used here in this problem. i.e. explicit temporary, bytemap format, transpositions}.
\begin{figure}
    \label{fig:spgemm_small}
	\includegraphics[width=\linewidth]{spgemm_small_speedup_log_scale.png}
    \caption{A comparison of several matrix multiplication algorithms between Finch and Taco smaller matrices, ordered from small to big dimension. Note that inner products necessarily requires $O(n^2)$ work and taco's outer products format is dense. Finch can use a sparse outer products format and thus has an asymptotic advantage that becomes evident as the output dimensions grow.}
\end{figure}

\begin{figure}
\label{fig:spgemm_large}
	\includegraphics[width=\linewidth]{spgemm_joel_speedup.png}
    \caption{A comparison of gustavson's algorithm between Finch and Taco on some larger matrices}
\end{figure}
